---
output:
  md_document:
    variant: markdown_github
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

bioenergmod
===============

The bioenergmod package contains a set of functions developed for bioenergetics modeling, to compare the daily energy requirements of a wildlife population of interest to the daily energy supply available, taking into account dynamic habitat availability and consumption in previous time steps. This was primarily developed with wetland-dependent shorebirds in mind, to account for variation in the proportion of potential habitat that is flooded (i.e. has open water) and variation in the proportion of the open water that is accessible (i.e. shallow enough for use by foraging shorebirds).

Features
--------------
* Accommodates multiple land cover types with differing availability and foraging value
* Accommodates tracking energy supply in all available habitat vs. accessible habitat (e.g. if only a proportion of available habitat is also accessible)
* Includes option for Monte Carlo simulation
* Includes plotting function

Installation
---------------
You can install this package directly from R using the **devtools** package:

```{r, eval=FALSE}
devtools::install_github("kdybala/bioenergmod")
```

References
---------------
For details on the model's structure:
Dybala KE, Reiter ME, Hickey CM, Shuford WD, Strum KM, Yarris GS. <i>In review</i>. A bioenergetics modeling approach to setting conservation objectives for non-breeding shorebirds in California's Central Valley. 

Example
---------------
```{r Example}
library(bioenergmod)

## Calculate daily energy requirements for a shorebird from population size (n), average individual body mass (kg), and assimilation efficiency.
energy.need = calculate_energy_demand(n=c(5000,7000,10000), bodymass=c(0.1,0.08,0.05), metabolism='FMR', assimilation=0.73)

## Habitat foraging value (kJ/ha):
energy.density = data.frame(habitat=c('A','B','C'), value=c(20,15,12))

## Habitat availability:
total.area = data.frame(habitat=c('A','B','C'), area=c(70000,150000,100000))
prop.openwater = data.frame(habitat=c(rep('A',3),rep('B',3),rep('C',3)), yday=rep(c(1:3),3), 
                            value=c(0.9,0.85,0.95, 0.2,0.5,0.8, 0.5,0.9,0.7))
prop.accessible = data.frame(habitat=c(rep('A',3),rep('B',3),rep('C',3)), yday=rep(c(1:3),3), 
                             value=rep(0.9,9))

## Calculate daily change in habitat availability:
change = calculate_habitat_change(tothabitat=total.area, flood=prop.openwater, time='yday', value='value', accessible=prop.accessible, wetsplit=F)

## Run bioenergetics model:
results = run_bioenergmod_loop(energyneed=energy.need, energydens=energy.density, 
                               habitat.available=change$openwater, habitat.accessible=change$accessible,
                               habitat.added=change$added, habitat.returned=change$returned,
                               prop.accessible=change$prop.accessible)

## Plot accessible energy by time step
plot_bioenergmod(results$energy.accessible, scale=1000000, ylab='kJ (millions)', der=energy.need)

## Plot shortfalls in meeting daily energy requirements:
library(ggplot2)
ggplot(results$energy, aes(x=time, y=shortfall/1000000)) + geom_area(fill='gray80') + ylab('kJ (millions)') + theme_classic()

```


